Data Prep

### 2024-01-13 00.0_dataPrep.qmd
finalized cleaning up the data pipeline

## Goals and introduction

@@

## Setup

@@

@@

# tidy up
rm(ornitela)
gc()
```

@@

@@

-   `removed`

@@

## Remove capture cage periods

@@

@@

# tidy up
rm(AllCarmelDates)
rm(AllCarmelDates.1)
rm(AllCarmelDates.2)
rm(AllCarmelDates.3)
rm(AllCarmelDates.all)
rm(capture_sites)
rm(crds)
rm(DistanceMat)
rm(removed)
rm(removed_captures)
rm(roosts0)
rm(sub.captured)
rm(sub.captured.carmel)
rm(sub.captured.dates)
rm(sub.captured.dates.1)
rm(sub.captured.no.carmel)
rm(sub.roosts)
rm(removed_periods)
rm(after)
rm(before)
rm(ClosestCaptureDist)
rm(ClosestCaptureSite)
rm(distance)
gc()
```

@@

START HERE!!!

@@

-   `data/derived/toRemove_lfr.Rda`

-   `data/derived/seasons_list_removed_lfr.Rda`

@@

# Exclude individuals that never have a daily mode of 10 minutes
toRemove_lfr <- map(with_modes, ~.x %>%
                        sf::st_drop_geometry() %>%
                        group_by(Nili_id) %>%
                        summarize(minmode = min(mode, na.rm = T)) %>%
                        filter(minmode > 10) %>%
                        pull(Nili_id) %>%
                        unique())
save(toRemove_lfr, file = "data/derived/toRemove_lfr.Rda")
seasons_list_removed_lfr <- map2(seasons_list, toRemove_lfr, ~.x %>% filter(!(Nili_id %in% .y)))
save(seasons_list_removed_lfr, file = "data/derived/seasons_list_removed_lfr.Rda")

# Tidy up
rm(seasons_list)
rm(seasons_list_removed_lfr)
rm(with_modes)
rm(toRemove_lfr)
gc()
```

@@

The next step will be to remove individuals with too few points per day or too few days tracked. But we don't want those indivs removed for the *social* analysis, since those things will be accounted for with SRI and all individuals make up important parts of the social fabric. So when we do the social analysis, we will use `seasons_list_removed_lfr` and not anything further along.

@@

-   `data/derived/seasons_list_removed_lfr`

@@

-   \`data/derived/roosts_seasons_list.

```{r}
load("data/derived/seasons_list_removed_lfr.Rda")
```

```{r}
#| eval: false
# Get roosts for each season ----------------------------------------------
roosts <- purrr::map(seasons_list_removed_lfr, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id"))
roosts <- roosts %>%
  map(., ~st_as_sf(.x, crs = "WGS84", 
                   coords = c("location_long", "location_lat"), 
                   remove = F))
save(roosts, file = "data/derived/roosts.Rda")

# Tidy up
rm(seasons_list_removed_lfr)
rm(roosts)
gc()
```

@@

START HERE

XXX do we not also remove nighttime points for seasons_mode10?

@@

-   `seasons`

@@

-   `seasons`

```{r}
#| eval: false
# Remove nighttime points -------------------------------------------------
before <- map_dbl(seasons, nrow)
seasons <- map(seasons, ~{
  times <- suncalc::getSunlightTimes(date = unique(lubridate::date(.x$timestamp)),
                                     lat = 31.434306, lon = 34.991889, # XXX revisit this!
                                     keep = c("sunrise", "sunset")) %>%
    dplyr::select("dateOnly" = date, sunrise, sunset)

  .x <- .x %>%
    dplyr::mutate(dateOnly = lubridate::ymd(dateOnly)) %>%
    dplyr::left_join(times, by = "dateOnly") %>%
    dplyr::mutate(daylight = ifelse(timestamp >= sunrise & timestamp <= sunset, "day", "night")) %>%
    dplyr::select(-c(sunrise, sunset)) %>%
    dplyr::filter(daylight == "day")
})
after <- map_dbl(seasons, nrow)
change <- after-before
changeDF <- data.frame(seasonName = seasonNames_orig, nBefore = before, nAfter = after) %>%
  mutate(propRowsRemoved = round((before-after)/before, 2))
changeDF
```

## Restrict to Southern individuals

@@

XXX why are we using `seasons_orig` here? shouldn't it be `seasons`?

@@

-   `seasons_orig`

@@

-   `data/derived/seasonsSF.Rda`

-   `seasons` XXX seems wrong!

```{r}
#| eval: false
# Restrict to southern individuals ----------------------------------------
# 
# Transform to SF object, so we can get centroids
seasonsSF <- map(seasons_orig, ~.x %>%
                   sf::st_as_sf(coords = c("location_long", "location_lat"), remove = F) %>%
                   sf::st_set_crs("WGS84") %>%
                   sf::st_transform(32636))

save(seasonsSF, file = "data/derived/seasonsSF.Rda")
load("data/derived/seasonsSF.Rda")

## Get centroids, so we can see who's "southern" for that season.
centroids <- map(seasonsSF, ~.x %>%
                   group_by(Nili_id) %>%
                   summarize(geometry = st_union(geometry)) %>%
                   st_centroid())

@@

## Get southern individuals for each season, so we can filter the data
southernIndivs <- map(centroids, ~.x %>%
                        filter(st_coordinates(.)[,2] < 3550000) %>%
                        pull(Nili_id))

## Remove individuals not in the south
before <- seasons
seasons <- map2(.x = seasons, .y = southernIndivs, ~.x %>% filter(Nili_id %in% .y))
after <- seasons

beforeRows <- map_dbl(before, nrow)
afterRows <- map_dbl(after, nrow)
beforeIndivs <- map_dbl(before, ~length(unique(.x$Nili_id)))
afterIndivs <- map_dbl(after, ~length(unique(.x$Nili_id)))
df <- data.frame(season = seasonNames_orig, nBefore = beforeRows, nAfter = afterRows, indivsBefore = beforeIndivs, indivsAfter = afterIndivs)
df$propRowsRemoved <- round((df$nBefore - df$nAfter)/df$nBefore, 2)
df
rm(centroids)
rm(seasonsSF)
```

@@

-   `seasons`

@@

-   `seasons`

```{r}
#| eval: false
# Include only individuals with enough points per day ------------------------
# Investigate battery percentage vs points per day ------------------------
# Is there some way we can relate battery charge percentage to number of points per day?
test <- seasons[[5]]
names(test)

ppd <- map_dfr(seasons, ~.x %>%
                 st_drop_geometry() %>%
                 group_by(seasonUnique, season, Nili_id, dateOnly) %>%
                 summarize(n = n(),
                           minBatt = min(battery_charge_percent, na.rm = T),
                           meanBatt = mean(battery_charge_percent, na.rm = T),
                           medBatt = mean(battery_charge_percent, na.rm = T)))

@@

# # cumulative histogram to look at the results of removing days based only on ppd
# ppd %>%
#   filter(n < 100) %>%
#   ggplot(aes(x = n)) +
#   stat_ecdf(aes(y = after_stat(y)*100))+
#   ylab("Percentage of vulture-days included")+
#   xlab("Points per day")+
#   theme_classic()+
#   facet_wrap(~seasonUnique)
# 
# # cumulative histogram to look at results of removing days based on ppd after battery is taken into account (min pct < 50)
# ppd %>%
#   filter(n < 100, minBatt > 50) %>%
#   ggplot(aes(x = n)) +
#   stat_ecdf(aes(y = after_stat(y)*100))+
#   ylab("Percentage of vulture-days included")+
#   xlab("Points per day")+
#   theme_classic()+
#   facet_wrap(~seasonUnique)
# 
# # same thing, but with batery set at 75
# # cumulative histogram to look at results of removing days based on ppd after battery is taken into account (min pct < 50)
# ppd %>%
#   filter(n < 100, minBatt > 75) %>%
#   ggplot(aes(x = n)) +
#   stat_ecdf(aes(y = after_stat(y)*100))+
#   ylab("Percentage of vulture-days included")+
#   xlab("Points per day")+
#   theme_classic()+
#   facet_wrap(~seasonUnique)
# 
# # What if instead we decide we want a certain number of ppd--battery cuttoff may be sharper. Say, 30 ppd: is there a discernible battery cutoff that distinguishes "real" from "battery artifact" days?
# ppd %>%
#   filter(n < 30) %>%
#   ggplot(aes(x = minBatt))+
#   stat_ecdf(aes(y = after_stat(y)*100))+
#   ylim(c(0, 100))+
#   facet_wrap(~seasonUnique)+
#   ylab("Percentage of vulture-days included")+
#   xlab("Minimum battery percentage") # we do seem to see crooks here around 75%!
# 
# # Same thing but with 10 points per day? Predict much sharper cutoffs.
# ppd %>%
#   filter(n < 10) %>%
#   ggplot(aes(x = minBatt))+
#   stat_ecdf(aes(y = ..y..*100))+
#   ylim(c(0, 100))+
#   facet_wrap(~seasonUnique)+
#   ylab("Percentage of vulture-days included")+
#   xlab("Minimum battery percentage") # crooks around 50% battery minimum for 10ppd. This seems like a pretty good way to go!

# Let's throw out anything with < 10ppd unless the battery has a min charge > 50%.

# Orr suggests to just restrict to vulture*days with >30 points per day. Let's see what that would look like.
thresh <- 30
# ppd %>%
#   mutate(Status = ifelse(n < thresh, "removed", "kept")) %>%
#   ggplot(aes(x = n))+
#   geom_histogram(aes(fill = Status))+
#   facet_wrap(~seasonUnique, scales ="free")+
#   geom_vline(aes(xintercept = thresh), col = "red")+
#   scale_fill_manual(values = c("darkgray", "red"))+
#   theme_classic()+
#   ggtitle(paste("n removed =", thresh, sep = " "))+
#   ylab("# vulture*days")+
#   xlab("# points per vulture*day") # short days in breeding season, and vultures sitting on nests
# 
# ppd <- ppd %>%
#   ungroup() %>%
#   mutate(removed = ifelse(n < thresh, T, F))
# 
# # what proportion of vulture*days are removed overall?
# (propRemoved <- ppd %>%
#     group_by(seasonUnique, season) %>%
#     summarize(nTot = n(),
#               nRemoved = sum(removed),
#               propRemoved = nRemoved/nTot))
# propRemoved %>%
#   ggplot(aes(x = seasonUnique, y = propRemoved, col = season))+
#   geom_point(size = 10)+
#   theme_classic()+
#   scale_color_manual(values = c("blue", "red", "darkorange"))
# 
# # I'm a bit concerned about this because it removes more data for the breeding season than for the non-breeding seasons (likely because less sunlight during those seasons)
# 
# # what about by individual over each season?
# ppd %>%
#   group_by(seasonUnique, season, Nili_id) %>%
#   summarize(propRemoved = sum(removed)/n()) %>%
#   ggplot(aes(x = seasonUnique, y = propRemoved))+
#   geom_boxplot(aes(group = seasonUnique, fill = season))+
#   geom_jitter(width = 0.2, alpha = 0.3, size = 2)+
#   scale_fill_manual(name = "Season", values = c("blue", "red", "darkorange"))+
#   theme_classic()+ # tells the same story
#   ylab("Proportion of days removed")+
#   xlab("")

# Based on the battery charge analysis above, going to remove individuals that have 10 or fewer points per day, but only if the minimum battery charge is less than 50%.

## Keep days with more than 10 points, or with fewer than 10 if the minimum battery charge that day is >50%
seasons <- map(seasons, ~.x %>%
                 group_by(Nili_id, dateOnly) %>%
                 filter(n() >= 10 | (n() < 10 & min(battery_charge_percent, na.rm = T) > 50)) %>%
                 ungroup())
```

@@

@@

-   `seasons`

@@

-   `seasons`

```{r}
#| eval: false
# Include only individuals with enough days tracked -----------------------
## Must be tracked for at least 30 days per season
durs <- map_dbl(seasons, ~length(unique(.x$dateOnly)))

# walk2(seasons, durs, ~.x %>%
#         st_drop_geometry() %>%
#         group_by(Nili_id) %>%
#         summarize(nDaysTracked = length(unique(dateOnly)),
#                   propDaysTracked = nDaysTracked/.y) %>%
#         pull(propDaysTracked) %>%
#         hist())

indivsToKeep <- map2(.x = seasons, .y = durs, ~.x %>%
                       st_drop_geometry() %>%
                       group_by(Nili_id) %>%
                       summarize(nDaysTracked = length(unique(dateOnly)),
                                 propDaysTracked = nDaysTracked/.y) %>%
                       filter(nDaysTracked >= 30) %>%
                       pull(Nili_id))

## remove those individuals not tracked for enough days
before <- seasons
after <- map2(.x = seasons, .y = indivsToKeep, ~.x %>% filter(Nili_id %in% .y))

beforeIndivs <- map_dbl(before, ~length(unique(.x$Nili_id)))
afterIndivs <- map_dbl(after, ~length(unique(.x$Nili_id)))
seasons <- after
```

@@

-   `seasons`

@@

-   `data/derived/elevs_z10_seasons.Rda`

-   `data/derived/groundElev_z10.Rda`

-   `seasons`

```{r}
#| eval: false
# Get elevation rasters ---------------------------------------------------
# Remove the first element of `seasons` because it doesn't have any data
seasons <- seasons[-1] # removing summer 2020
# # XXXXXXX start here
#seasons <- map(seasons, ~st_as_sf(.x, coords = c("location_lat", "location_long"), remove = F, crs = "WGS84"))
#elevs_z10_seasons <- map(seasons, ~elevatr::get_elev_raster(.x , z = 10))
#save(elevs_z10_seasons, file = "data/derived/elevs_z10_seasons.Rda")
load("data/derived/elevs_z10_seasons.Rda")

# groundElev_z10 <- map2(elevs_z10_seasons, seasons, ~raster::extract(x = .x, y = .y)) # have to convert the crs because the raster from elevatr is in WGS84.
# save(groundElev_z10, file = "data/derived/groundElev_z10.Rda")
load("data/derived/groundElev_z10.Rda")

## use elevation rasters to calculate height above ground level
before <- seasons
seasons <- map2(.x = seasons, .y = groundElev_z10, 
                ~.x %>% mutate(groundElev = .y) %>%
                  mutate(height_above_ground = height_above_msl-groundElev,
                         height_above_ground = case_when(height_above_ground < 0 ~ 0,
                                                         TRUE ~ height_above_ground)))
after <- seasons

# There should be no change in the number of rows
all(map2_lgl(before, after, ~nrow(.x) == nrow(.y)))
```

## Restrict to mode10 individuals

XXX what is happening here

@@

-   `data/derived/toKeep_fixrate.Rda`

-   `seasons`

@@

-   `data/derived/seasons.Rda`

-   `data/derived/seasons_mode10.Rda`

```{r}
#| eval: false
# Now restrict to the mode10 individuals ----------------------------------
load("data/derived/toKeep_fixrate.Rda")
toKeep_fixrate <- toKeep_fixrate[-1]
seasons_mode10 <- map2(seasons, toKeep_fixrate, ~.x %>% filter(Nili_id %in% .y))

save(seasons, file = "data/derived/seasons.Rda")
save(seasons_mode10, file = "data/derived/seasons_mode10.Rda")
load("data/derived/seasons.Rda")
load("data/derived/seasons_mode10.Rda")
```

Downsampling data to 10-minute intervals exclusively

-   `seasons`

-   `seasons_mode10`

-   `data/derived/seasons_10min.Rda`

-   `data/derived/seasons_mode10_10min.Rda`

XXX this is confusing. Need to rename after concepts, not minutes.

```{r}
#| eval: false
# Downsample the data and save the downsampled data
subsample <- function(df, idCol = "Nili_id", timestampCol = "timestamp", mins = 10){
  sub <- df %>%
    arrange(idCol, timestampCol) %>%
    group_by(.data[[idCol]]) %>%
    mutate(tc = cut(.data[[timestampCol]], 
                    breaks = paste(as.character(mins), "min", sep = " "))) %>%
    group_by(.data[[idCol]], 
             "d" = lubridate::date(.data[[timestampCol]]), tc) %>%
    slice(1) %>%
    ungroup() %>%
    dplyr::select(-c("d", "tc"))
  return(sub)
}

# create and save downsampled datasets for later use
seasons_10min <- map(seasons, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 10), .progress = T) # basically just trying to remove any data from the occasional bursts of more concentrated and frequent sampling.
seasons_mode10_10min <- map(seasons_mode10, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 10), .progress = T)

map_dbl(seasons, nrow)
map_dbl(seasons_10min, nrow)
map_dbl(seasons_mode10, nrow)

seasons_mode10_10min <- map(seasons_mode10_10min, ~.x %>% group_by(Nili_id) %>% mutate(daysTracked = length(unique(dateOnly))))
seasons_10min <- map(seasons_10min, ~.x %>% group_by(Nili_id) %>% mutate(daysTracked = length(unique(dateOnly))))

save(seasons_10min, file = "data/derived/seasons_10min.Rda")
save(seasons_mode10_10min, file = "data/derived/seasons_mode10_10min.Rda")
```

>>>>

# Goals and introduction

@@

# Setup

@@

# Processing for all data

@@

# tidy up
rm(ornitela)
rm(minDate)
rm(maxDate)
gc()
```

@@

### Hospital/pre-determined invalid

@@

@@

### Capture cages

@@

-   `data/raw/all_captures_carmel_2010-2021.csv`

@@

# tidy up
rm(AllCarmelDates)
rm(AllCarmelDates.1)
rm(AllCarmelDates.2)
rm(AllCarmelDates.3)
rm(AllCarmelDates.all)
rm(capture_sites)
rm(crds)
rm(DistanceMat)
rm(removed)
rm(removed_captures)
rm(roosts0)
rm(sub.captured)
rm(sub.captured.carmel)
rm(sub.captured.dates)
rm(sub.captured.dates.1)
rm(sub.captured.no.carmel)
rm(sub.roosts)
rm(removed_periods)
rm(after)
rm(before)
rm(ClosestCaptureDist)
rm(ClosestCaptureSite)
rm(distance)
rm(start.day)
rm(start.month)
rm(end.day)
rm(end.month)
rm(propChange)

gc()
```

@@

@@

-   `data/derived/removed_lfr.Rda`

@@

# Exclude individuals that never have a daily mode of 10 minutes
toRemove_lfr <- map(with_modes, ~.x %>%
                        sf::st_drop_geometry() %>%
                        group_by(Nili_id) %>%
                        summarize(minmode = min(mode, na.rm = T)) %>%
                        filter(minmode > 10) %>%
                        pull(Nili_id) %>%
                        unique())

removed_lfr <- map2(seasons_list, toRemove_lfr, ~.x %>% 
                                   filter(!(Nili_id %in% .y)))

save(removed_lfr, 
     file = "data/derived/removed_lfr.Rda")

# Tidy up
rm(seasons_list)
rm(removed_lfr)
rm(with_modes)
rm(toRemove_lfr)
gc()
```

@@

The next step will be to remove individuals with too few points per day or too few days tracked. But we don't want those indivs removed for the *social* analysis, since those things will be accounted for with SRI and all individuals make up important parts of the social fabric. So when we do the social analysis, we will use `removed_lfr` and not anything further along.

@@

-   `data/derived/removed_lfr`

@@

-   `data/derived/roosts.Rda`

```{r}
load("data/derived/removed_lfr.Rda")
```

```{r}
#| eval: false
# Get roosts for each season ----------------------------------------------
roosts <- purrr::map(removed_lfr, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id"))
roosts <- roosts %>%
  map(., ~st_as_sf(.x, crs = "WGS84", 
                   coords = c("location_long", "location_lat"), 
                   remove = F))
save(roosts, file = "data/derived/roosts.Rda")

# Tidy up
rm(removed_lfr)
rm(roosts)
gc()
```

# Further processing for movement data

@@

@@

-   `data/derived/removed_lfr.Rda`
-   `data/derived/season_names.Rda`

@@

-   `data/derived/removed_nightime.Rda`

```{r}
load("data/derived/removed_lfr.Rda")
load("data/derived/season_names.Rda")
```

```{r}
#| eval: false
# Remove nighttime points -------------------------------------------------
removed_nighttime <- map(removed_lfr, ~{
  times <- suncalc::getSunlightTimes(date = unique(lubridate::date(.x$timestamp)),
                                     lat = 31.434306, lon = 34.991889, # XXX revisit this!
                                     keep = c("sunrise", "sunset")) %>%
    dplyr::select("dateOnly" = date, sunrise, sunset)

  .x <- .x %>%
    dplyr::mutate(dateOnly = lubridate::ymd(dateOnly)) %>%
    dplyr::left_join(times, by = "dateOnly") %>%
    dplyr::mutate(daylight = ifelse(timestamp >= sunrise & timestamp <= sunset, "day", "night")) %>%
    dplyr::select(-c(sunrise, sunset)) %>%
    dplyr::filter(daylight == "day")
})

save(removed_nighttime, file = "data/derived/removed_nighttime.Rda")

# tidy up
rm(removed_nighttime)
rm(removed_lfr)
rm(season_names)
gc()
```

## Restrict to southern population

@@

@@

-   `data/derived/removed_nighttime.Rda`
-   `data/derived/season_names.Rda`

@@

-   `data/derived/removed_northern.Rda`

```{r}
load("data/derived/removed_nighttime.Rda")
load("data/derived/season_names.Rda")
```

```{r}
#| eval: false
# Transform to SF object, so we can get centroids
seasons_sf <- map(removed_nighttime, ~.x %>%
                   sf::st_as_sf(coords = c("location_long", "location_lat"), remove = F) %>%
                   sf::st_set_crs("WGS84") %>%
                   sf::st_transform(32636))

## Get centroids, so we can see who's "southern" for that season.
centroids <- map(seasons_sf, ~.x %>%
                   group_by(Nili_id) %>%
                   summarize(geometry = st_union(geometry)) %>%
                   st_centroid())

@@

## Get southern individuals for each season, so we can filter the data
southern_indivs <- map(centroids, ~.x %>%
                        filter(st_coordinates(.)[,2] < 3550000) %>%
                        pull(Nili_id))

## Remove individuals not in the south
before <- removed_nighttime
removed_northern <- map2(.x = removed_nighttime, 
                         .y = southern_indivs, ~.x 
                         %>% filter(Nili_id %in% .y))
after <- removed_northern

beforeRows <- map_dbl(before, nrow)
afterRows <- map_dbl(after, nrow)
beforeIndivs <- map_dbl(before, ~length(unique(.x$Nili_id)))
afterIndivs <- map_dbl(after, ~length(unique(.x$Nili_id)))
df <- data.frame(season = season_names, nBefore = beforeRows, nAfter = afterRows, indivsBefore = beforeIndivs, indivsAfter = afterIndivs)
df$propRowsRemoved <- round((df$nBefore - df$nAfter)/df$nBefore, 2)
df

save(removed_northern, file = "data/derived/removed_northern.Rda")

# tidy up
rm(removed_nighttime)
rm(season_names)
rm(seasons_sf)
rm(centroids)
rm(southern_indivs)
rm(before)
rm(after)
rm(removed_northern)
rm(df)
rm(afterIndivs)
rm(afterRows)
rm(beforeIndivs)
rm(beforeRows)
gc()
```

@@

-   `data/derived/removed_northern.Rda`

@@

-   `data/derived/removed_lowppd.Rda`

```{r}
load("data/derived/removed_northern.Rda")
```

```{r}
#| eval: false
# Calculate number of points per day for each individual
ppd <- map_dfr(removed_northern, ~.x %>%
                 st_drop_geometry() %>%
                 group_by(seasonUnique, season, Nili_id, dateOnly) %>%
                 summarize(n = n(),
                           minBatt = min(battery_charge_percent, na.rm = T),
                           meanBatt = mean(battery_charge_percent, na.rm = T),
                           medBatt = mean(battery_charge_percent, na.rm = T)))

# Sometimes, fewer points per day is due to low battery charge percent. Let's examine the relationship between battery charge and number of points per day.
ppd %>%
  filter(n < 100) %>% # only interested in lower numbers of points per day
  ggplot(aes(x = minBatt, y = n))+
  geom_point()+
  facet_wrap(~seasonUnique)+
  theme_classic()

# What if we decide to exclude based on the combination of battery charge and points per day, with the minimum battery threshold at 50% and the points threshold at 10?

# That is, if the minimum battery threshold was <= 50, a bird must have > 10 points on a given day for that day to be included. But if the minimum battery threshold was never less than 50%, it's okay to keep the ones with fewer points per day, since we know it was due to some reason other than battery charge.

@@

## Keep days with more than 10 points, or with fewer than 10 if the minimum battery charge that day is >50%
removed_lowppd <- map(removed_northern, ~.x %>%
                 group_by(Nili_id, dateOnly) %>%
                 filter(n() >= 10 | (n() < 10 & 
                                       min(battery_charge_percent, 
                                           na.rm = T) > 50)) %>%
                 ungroup())

save(removed_lowppd, file = "data/derived/removed_lowppd.Rda")

# tidying up
rm(removed_northern)
rm(ppd)
rm(excl_batt)
rm(excl_pts)
rm(removed_lowppd)
gc()
```

@@

Now that we've removed days with too few points per day, we are left with only "valid" movement days for each individual. Now we want to exclude *individuals* (rather than individual-days) if the individual has too few valid days of movement data to accurately calculate movement metrics.

Somewhat arbitrarily, we are removing individuals that are tracked for **fewer than 30 days in a given season.**

@@

-   `data/derived/removed_lowppd.Rda`
-   `data/derived/season_names.Rda`

@@

-   `data/derived/removed_too_few_days.Rda`
-   `data/derived/season_durations.Rda`

```{r}
load("data/derived/removed_lowppd.Rda")
load("data/derived/season_names.Rda")
```

```{r}
#| eval: false

# How long are the seasons?
season_durations <- map_dbl(removed_lowppd, ~max(.x$dateOnly)-min(.x$dateOnly))
names(season_durations) <- season_names
season_durations

# Include only individuals with enough days tracked -----------------------
indivsToKeep <- map(removed_lowppd, ~.x %>%
                       st_drop_geometry() %>%
                       group_by(Nili_id) %>%
                       summarize(nDaysTracked = 
                                   length(unique(dateOnly))) %>%
                       filter(nDaysTracked >= 30) %>%
                       pull(Nili_id))

## remove those individuals not tracked for enough days
removed_too_few_days <- map2(.x = removed_lowppd, 
                             .y = indivsToKeep, ~.x %>% 
                filter(Nili_id %in% .y))

beforeIndivs <- map_dbl(removed_lowppd, ~length(unique(.x$Nili_id)))
afterIndivs <- map_dbl(removed_too_few_days, ~length(unique(.x$Nili_id)))

save(removed_too_few_days, file = "data/derived/removed_too_few_days.Rda")
save(season_durations, file = "data/derived/season_durations.Rda")

# tidy up
rm(removed_lowppd)
rm(season_names)
rm(season_durations)
rm(indivsToKeep)
rm(removed_too_few_days)
rm(beforeIndivs)
rm(afterIndivs)
gc()
```

@@

-   `data/derived/removed_too_few_days`

@@

-   `data/derived/elevs_z10.Rda`

-   `data/derived/with_altitudes.Rda`

```{r}
load("data/derived/removed_too_few_days.Rda")
```

```{r}
# Get elevation rasters ---------------------------------------------------
sf <- map(removed_too_few_days, ~st_as_sf(.x, coords = c("location_lat", "location_long"), remove = F, crs = "WGS84"))
elevs_z10 <- map(sf, ~elevatr::get_elev_raster(.x , z = 10))
save(elevs_z10, file = "data/derived/elevs_z10.Rda")
```

```{r}
load("data/derived/elevs_z10.Rda")
```

```{r}
#| eval: false
groundElev_z10 <- map2(elevs_z10, sf, ~raster::extract(x = .x, y = .y)) # have to convert the crs because the raster from elevatr is in WGS84.

## use elevation rasters to calculate height above ground level
with_altitudes <- map2(.x = removed_too_few_days, .y = groundElev_z10, 
                ~.x %>% mutate(groundElev = .y) %>%
                  mutate(height_above_ground = height_above_msl-groundElev,
                         height_above_ground = case_when(height_above_ground < 0 ~ 0,
                                                         TRUE ~ height_above_ground)))

# There should be no change in the number of rows
all(map2_lgl(removed_too_few_days, with_altitudes, ~nrow(.x) == nrow(.y)))

save(with_altitudes, file = "data/derived/with_altitudes.Rda")

# tidy up
rm(removed_too_few_days)
rm(sf)
rm(elevs_z10)
rm(groundElev_z10)
rm(with_altitudes)
```

We already removed individuals that had too low a fix rate. We are left with individuals that had a fix rate of 10 minutes or more frequent. But occasionally, there were periods of high-frequency sampling. This might mess up our movement calculations, so let's downsample the data so that any points that were sampled at a frequency much higher than 1 fix/10 minutes will be reduced to the same frequency as the others.

@@

-   `data/derived/with_altitudes.Rda`

@@

-   `data/derived/downsampled_10min.Rda`

```{r}
load("data/derived/with_altitudes.Rda")
load("data/derived/season_names.Rda")
```

```{r}
#| eval: false

# Function to downsample the data
subsample <- function(df, idCol = "Nili_id", 
                      timestampCol = "timestamp", mins = 10){
  sub <- df %>%
    arrange(idCol, timestampCol) %>%
    group_by(.data[[idCol]]) %>%
    mutate(tc = cut(.data[[timestampCol]], 
                    breaks = paste(as.character(mins), "min", sep = " "))) %>%
    group_by(.data[[idCol]], 
             "d" = lubridate::date(.data[[timestampCol]]), tc) %>%
    slice(1) %>%
    ungroup() %>%
    dplyr::select(-c("d", "tc"))
  return(sub)
}

# create and save downsampled datasets for later use
downsampled_10min <- map(with_altitudes, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 10), .progress = T) # basically just trying to remove any data from the occasional bursts of more concentrated and frequent sampling.

# Add information on # of days tracked for each individual
downsampled_10min <- map(downsampled_10min, ~.x %>% group_by(Nili_id) %>% 
                       mutate(daysTracked = length(unique(dateOnly))))

save(downsampled_10min, file = "data/derived/downsampled_10min.Rda")

map_dbl(with_altitudes, nrow) %>% setNames(season_names)
map_dbl(downsampled_10min, nrow) %>% setNames(season_names)

# tidying
rm(with_altitudes)
rm(downsampled_10min)
gc()
```
